---
title: "DEA analysis"
author: "Riksrevisjonen"
output: pdf_document
params:
  data: NA
  idvar: NA
  inputvars: NA
  outputvars: NA
  normdata: NA
  dearts: NA
  deaorient: NA
  deain: NA
  deaout: NA
  deanorm: NA
  modelout: NA
---

```{r, setup, include=FALSE}
mod <- list(
  rts = switch(params$dearts,
    'crs' = 'constant returns to scale',
    'vrs' = 'variable returns to scale',
    'nirs' = 'non-increasing returns to scale',
    'ndrs' = 'non-decreasing returns to scale'
  ),
  orient = switch(params$deaorient,
    'in' = 'input oriented',
    'out' = 'output oriented'
  )
)
```

## Dataset

The dataset consists of `r nrow(params$data)` observations and `r ncol(params$data)` variables.

## Variables

The following variables were used as input variables: `r paste(params$inputvars, sep = ', ')`.

```{r, echo=FALSE}
df <- data.frame(
  min = sapply(params$inputvars, function(x) min(params$data[, x])),
  q25 = sapply(params$inputvars, function(x) quantile(params$data[, x])[[2]]),
  median = sapply(params$inputvars, function(x) median(params$data[, x])),
  q75 = sapply(params$inputvars, function(x) quantile(params$data[, x])[[4]]),
  max = sapply(params$inputvars, function(x) max(params$data[, x])),
  mean = sapply(params$inputvars, function(x) mean(params$data[, x]))
)
colnames(df) <- c('Min.', '1st Qu.', 'Median', '3rd. Qu.', 'Max', 'Mean')
knitr::kable(df)
```

The following variables were used as output variables: `r paste(params$outputvars, sep = ', ')`.

```{r, echo=FALSE}
df <- data.frame(
  min = sapply(params$outputvars, function(x) min(params$data[, x])),
  q25 = sapply(params$outputvars, function(x) quantile(params$data[, x])[[2]]),
  median = sapply(params$outputvars, function(x) median(params$data[, x])),
  q75 = sapply(params$outputvars, function(x) quantile(params$data[, x])[[4]]),
  max = sapply(params$outputvars, function(x) max(params$data[, x])),
  mean = sapply(params$outputvars, function(x) mean(params$data[, x]))
)
colnames(df) <- c('Min.', '1st Qu.', 'Median', '3rd. Qu.', 'Max', 'Mean')
knitr::kable(df)
```

## Model summary

Technology is `r mod$rts` and orientation is `r mod$orient`.

```{r, echo=FALSE}
eps <- 1e-06
eff <- params$modelout$eff

if (params$modelout$ORIENTATION != 'out' && is.null(params$modelout$direct)) {
  
  minE <- floor(10 * min(eff))/10
  dec <- seq(from = minE, to = 1, by = 0.1)
  
  estr <- sapply(1:length(dec), function(i) {
    if (i < length(dec))
      paste(dec[i], '<= E <', dec[i + 1])
    else if (i == length(dec))
      "E == 1"
  })
  
  num <- sapply(1:length(dec), function(i) {
    if (i < length(dec))
      sum(dec[i] - eps <= eff & eff < dec[i + 1] - eps)
    else if (i == length(dec))
      sum(abs(eff - 1) < eps)
  })
  
} else if (is.null(params$modelout$direct)) {
  
  maxF <- ceiling(10 * max(eff))/10
  dec <- seq(from = 1, to = maxF, by = 0.1)
  if (length(dec) > 10) {
    dec_ <- c(1, 1.1, 1.2, 1.3, 1.5, 2, 5, 10, 100, Inf)
    dec <- dec_[1:(max(which(dec_ < maxF)) + 1)]
  }
  
  estr <- sapply(1:length(dec), function(i) {
    if (i == 1)
      "F == 1"
    else if (i > 1)
      paste(dec[i - 1], '< F =<', dec[i])
  })
  
  num <- hist(eff, breaks = dec, plot = FALSE)$counts
  num[1] <- num[1] - sum(abs(eff - 1) < eps)
  num <- c(sum(abs(eff - 1) < eps), num)
  
} else {
  
  return(NULL)
  
}

eff.tbl <- data.frame(eff = estr, n = num, stringsAsFactors = FALSE)
colnames(eff.tbl) <- c('Efficiency range', 'Number of observations')

sum.tbl <- data.frame(
  min = min(eff), p25 = quantile(eff)[[2]], p50 = median(eff),
  m = mean(eff), p75 = quantile(eff)[[4]], max = max(eff)
)
colnames(sum.tbl) <- c('Min.', '1st Qu.', 'Median', 'Mean', '3rd. Qu.', 'Max')

knitr::kable(eff.tbl)
knitr::kable(sum.tbl)


# list(
#   p(class = 'h5', 'Summary of DEA analysis'),
#   p(paste('Technology is', params$modelout$RTS, 'and', params$modelout$ORIENTATION)),
#   p(paste('Mean efficiency:', round(mean(eff), 2))),
#   renderTable({ eff.tbl }),
#   renderTable({ sum.tbl }),
#   renderPlot({ hist(eff, col = 'red') })
# )
```

## Efficiency scores

The efficiency scores are estimated with the `Benchmarking` package in R. Based on the input variables, output variables, returns to scale and orientation stated above, the following efficiency scores has been calculated:

```{r, echo=FALSE}
d <- data.frame(`Efficiency score` = params$modelout$eff)
knitr::kable(d)
```

## Distribution

```{r, eval=TRUE, echo=FALSE}
if (params$deanorm) {
  txt <- 'The input data has been normalized so that each variable has a mean of 1.'
} else {
  txt <- 'The input data has not been altered.'
}
```

Below is a Salter diagram of all the decision making units in the analysis. The height of the bars are determined by the efficiency score for each unit. The width of the bars are determined by the sum of the combined inputs. `r txt`

```{r, eval=params$normdata, echo=FALSE, results='asis'}
# This will be printed as markdown if `params$normdata` is `TRUE`
# Remember to escape slashes!
cat(paste(
  "Normalizing data is done in two steps. First the column mean is calculated for",
  "each column in the input and output matrices respectively:"
))
cat("$$\\bar{V}_i=\\frac{\\sum_{n-1}^N V_{ni}}{N}$$")
cat(paste(
  "Where $\\bar{V}_i$ is the mean value for column $i$ (an input or output), $N$ is",
  "the number of DMUs and $V_{ni}$ is the value of DMU $n$ for a given input or output $i$."
))
cat("$$Vnorm_{ni}=\\frac{V_{ni}}{V_i}$$")
cat(paste(
  "Where $Vnorm_{ni}$ is the normalized value for the value associated with DMU $n$ and",
  "input or ouput in column $i$."
))
```

```{r, eval=TRUE, echo=FALSE}
require(ggplot2)

d <- params$data
d$eff <- params$modelout$eff

ins <- params$deain

if (is.matrix(ins) && dim(ins)[2] > 1) {
  w <- matrix(1, nrow = dim(ins)[2], ncol = 1)
  ins <- ins %*% w
}

d$ins <- ins
d <- d[order(d$eff), ]
d$right <- cumsum(d$ins)
d$left <- d$right - d$ins

ggplot(d, aes(ymin = 0)) +
  geom_rect(aes(xmin = left, xmax = right, ymax = eff)) +
  xlab('Combined inputs') + ylab('Efficiency score')
```

## Reproducability

The code below can be used to generate the DEA analysis.

The code assumes that our `data.frame` is available as the object `d`. You can download the data used in the analysis as an RDS-file. Read in the data with `d <- readRDS('dea.rds')`. The data must be in the current working directory in R. You can get the current working directory in R with `getwd()` and set the working directory to a new location with `setwd()`.

```{r, include=FALSE}
library(knitr)
library(stringr)
default_source_hook <- knit_hooks$get('source')

# Solution from
# https://stackoverflow.com/questions/43699235/replace-variables-with-their-corresponding-values-in-source-code-chunk

knit_hooks$set(source = function(x, options) {
  x <- str_replace_all(
    x, pattern = 'params\\$idvar', paste0("'", params$idvar, "'"))
  x <- str_replace_all(
    x, pattern = 'params\\$inputvars',
    paste0("'", paste(params$inputvars, collapse = '\',\n\  \''), "'"))
  x <- str_replace_all(
    x, pattern = 'params\\$outputvars',
    paste0("'", paste(params$outputvars, collapse = '\',\n\  \''), "'"))
  x <- str_replace_all(
    x, pattern = 'params\\$normdata', paste0("'", params$normdata, "'"))
  x <- str_replace_all(
    x, pattern = 'params\\$dearts', paste0("'", params$dearts, "'"))
  x <- str_replace_all(
    x, pattern = 'params\\$deaorient', paste0("'", params$deaorient, "'"))
  default_source_hook(x, options)
})
```

```{r, echo=TRUE, eval=FALSE}
# Note: you need to load the data set first as `d`
# Use `d <- readRDS('dea.rds')` to load the data
# Remember to check the working directory with `getwd()`

library(Benchmarking)

dea.id <- params$idvar
dea.in <- c(params$inputvars)
dea.out <- c(params$outputvars)
norm.data <- params$normdata

# Get the number of input variables
cx <- length(dea.in)

# We create a matrix with the values of all input variables
x <- matrix(
  sapply(dea.in, function(i) { d[[i]] }), ncol = cx,
  dimnames = list(d[,1], rep('x', cx))
)

# If `norm.data` is TRUE, we normalize the data in each column
if (norm.data == 'TRUE') {
  v <- colSums(x) / nrow(x)
  for (i in 1:ncol(x)) {
    x[,i] <- x[,i] / v[i]
  }
}

# Get the number of output variables
cy <- length(dea.out)

# We create a matrix with the values of all input variables
y <- matrix(
  sapply(dea.out, function(i) { d[[i]] }), ncol = cy,
  dimnames = list(d[,1], rep('y', cy))
)

if (norm.data == 'TRUE') {
  v <- colSums(y) / nrow(y)
  for (i in 1:ncol(y)) {
    y[,i] <- y[,i] / v[i]
  }
}

# Estimate the DEA model
dea.prod <- Benchmarking::dea(
  x, y, RTS = params$dearts, ORIENTATION = params$deaorient
)

# Print out the summary of the DEA model
summary(dea.prod)

# Estimate slack for all DMUs using the DEA model
slack <- Benchmarking::slack(x, y, dea.prod)

# Estimate super efficiency
super.eff <- Benchmarking::sdea(x, y, RTS = params$dearts, ORIENTATION = params$deaorient)
```
